### 2020-08-14 _meta/templates/Introduction.Rmd
refactor

---
title:
  Concrete compressive strength prediction with machine learning
date: "April 22, 2020"
author: "Pedro Bernardino Alves Moreira"
indent: true
abstract: "Compressive strength is the main characteristic of concrete. The correct prediction of this parameter means cost and time reduction. This work built predictive models for 6 different ages of concrete samples (3, 7, 14, 28, 56, and 100 days). A set of data obtained in previous studies was used, a total of 1030 samples, with 9 variables: compressive strength, age, and 7 ingredients (water, cement, fine aggregate, coarse aggregate, fly ash, blast furnace slag, and superplasticizers). Another 6 variables were added to represent the proportions of the main ingredients in each sample (water/cement, fine aggregate/cement, coarse aggregate/cement, fine aggregate/coarse aggregate, water/coarse aggregate, and water/fine aggregate). The predictive models were developed in *R* language, using the *caret* package with the *Parallel Random Forest* algorithm and repeated cross-validation technique to optimize the parameters. The results were satisfactory and compatible with other studies using the same data set. The most important model, 28 days old, obtained *RMSE* of 4.717. The 3-day model obtained the best result, *RMSE* of 3.310. The worst result was the 56-day model, with *RMSE* of 5.939. The work showed that the compressive strength of concrete can be predicted. The choice of creating a model for each age, instead of using age as a predictor, allowed to get compatible results with the available data at each age. It was a promising alternative since good results were achieved by training with just one algorithm. This work facilitates exploration and new efforts to predict the compressive strength of concrete, it can be replicated using different algorithms or the combination of several."
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
    toc_depth: 2
# bibliography: references.bib
references:
- id: Yeh1998
  author:
  - family: Yeh
    given: I-Cheng
  year: 1998
  month: 12
  pages: 1797-1808
  title: Modeling of Strength of High-Performance Concrete Using Artificial Neural Networks
  volume: 28
  journal: Cement and Concrete Research
  DOI: 10.1016/S0008-8846(98)00165-3
- id: Hameed2020
  author:
  - family: Hameed
    given: Mohammed
  - family: Khalid
    given: Mohamed
   year: 2020
   month: 01
   pages: 323-335
   title: "Prediction of Compressive Strength of High-Performance Concrete: Hybrid Artificial Intelligence Technique"
   # isbn: 978-3-030-38751-8
   DOI: 10.1007/978-3-030-38752-5_26
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{supertabular}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
- \usepackage{caption}
- \usepackage{floatrow}
- \floatsetup[table]{capposition=top}
- \floatsetup[figure]{capposition=top}
- \captionsetup{options=chunk}
- \DeclareNewFloatType{chunk}{placement=H, fileext=chk, name=}
- \renewcommand{\thechunk}{\arabic{chunk}}
- \usepackage{indentfirst}
- \usepackage{sectsty} \sectionfont{\centering}
---
```{r setup, include=FALSE, cache=F}
# Install libraries, if not already installed
if(!require(tidyverse)) install.packages("tidyverse")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(caret)) install.packages("caret")
if(!require(knitr)) install.packages("knitr")
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(gdata)) install.packages("gdata")
if(!require(dplyr)) install.packages("dplyr")
if(!require(tidyr)) install.packages("tidyr")
if(!require(cowplot)) install.packages("cowplot")
if(!require(pastecs)) install.packages("pastecs")
if(!require(factoextra)) install.packages("factoextra")
if(!require(reshape2)) install.packages("reshape2")
if(!require(purrr)) install.packages("purrr")
if(!require(gridExtra)) install.packages("gridExtra")
if(!require(questionr)) install.packages("questionr")
# Load libraries
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(gdata)
library(reshape2)
library(cowplot)
library(pastecs)
library(factoextra)
library(purrr)
library(gridExtra)
library(questionr)
# Set global options
options(pillar.sigfig = 5)
options(knitr.kable.NA = '')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo = TRUE, fig.align = 'center', cache=T, prompt=F, highlight=T )
oldSource <- knit_hooks$get("source")
knit_hooks$set(source = function(x, options) {
  x <- oldSource(x, options)
  x <- ifelse(!is.null(options$ref), paste0("\\label{", options$ref,"}", x), x)
  ifelse(!is.null(options$codecap), paste0("\\captionof{chunk}{", options$codecap,"}", x), x)
})
```

\newpage
\tableofcontents
\newpage
\listoffigures
\listoftables
\newpage

@@

# Materials and Methods

## Materials

The methodology was carried out using RStudio software [@RStudio], an integrated virtual environment for code development in *R* [@RCore]. Throughout the process, all code executed was documented in the same order as its execution in [Appendix 2](#appendix2), and reference was always made to codes throughout the text. All relevant information related to the operating system and installed packages has been presented in [Appendix 1](#appendix1).

## Reproducibility

In order to guarantee reproducibility, whenever there was some code that could use probabilistic operations, a *seed* was defined before its execution, ensuring that when run on another machine, with the same version of *R* and the same *seed*, get the same result. The *seeds* can be checked throughout [Appendix 2](#appendix2).

## Obtaining the data

The data was downloaded from the University of California Irvine website [@downloadData] (\ref{show-download-data}). In total there are 1030 samples with 9 columns. The samples were renamed and an id column was added to facilitate data manipulation (\ref{show-rename-dat-cols}). The columns were reordered to put the new id column in the first position (\ref{show-reorder-dat}). The first samples can be viewed in the table \ref{tab:first-samples}.

@@

```{r download-data, echo=F, include=F}
# Data download
# url_base <- "https://archive.ics.uci.edu"
# url <-  "/ml/machine-learning-databases/concrete/compressive/Concrete_Data.xls"
# download.file(paste0(url_base, url), "data.xls")
dat <- read.xls("data.xls")
n_inicial_samples <- nrow(dat)
colnames(dat)
```

```{r rename-dat-cols, echo=F}
# Renaming the columns
colnames(dat) <- c(
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
  )
dat$id <- seq.int(nrow(dat))
```

```{r reorder-dat, echo=F}
# Reordering the data
col_order <- c(
  "id",
  "cement",
  "blast_furnace_slag",
  "fly_ash",
  "water",
  "superplasticizers",
  "coarse_aggregate",
  "fine_aggregate",
  "day",
  "mpa"
)
dat <- dat[, col_order]
```

```{r col-names-and-units, echo=F}
# Defining column names and units
colNames <- c("ID", "Cement", "B.F.S.", "Fly ash", "Water",
             "Superp.", "C.Aggregate", "F.Aggregate", "Day", "Comp.Str.")
dfUnits <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
             "$kg/m^3$","$kg/m^3$","$kg/m^3$","","$MPa$")
```

```{r first-samples, echo=F}
# Table - First samples
caption <- "First 6 samples"
kable(
    dat[1:6,],
    col.names = dfUnits,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

## Data preparation

The preparation of the data consisted of transforming the sample set in order to maintain only relevant data for the subsequent studies. Data that were considered irrelevant or that had the potential to add undesirable noise to the analysis were removed. In addition, the relevant data has been transformed to better fit the studies in the next steps.

### Initial data cleaning

Initially, there were 25 duplicate samples that were removed, resulting in a new total of 1005 samples (\ref{show-removing-duplicated-samples}).

@@

```{r removing-duplicated-samples, echo=F}
# Removing duplicated samples
n_distinct_samples <- dat %>% select(-c(id)) %>% n_distinct()
n_duplicated_samples <- n_inicial_samples - n_distinct_samples
dat <- dat[!duplicated(select(dat, -c(id))),]
n_samples <- nrow(dat)
```

The data show the variables in the columns and samples in the rows. However it was found that some samples are identical in proportions of ingredients, changing only the value of age and compressive strength, for example, samples 653, 654, 678 and 681, shown in the table \ref{tab:similar-samples}.

@@

```{r similar-samples, echo=FALSE}
# Table - Samples with same composition
same_samples <- dat %>%
  filter(id %in% c(653, 678, 654, 681))
caption <- "Samples with same composition"
kable(
  same_samples[order(same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

In addition, there are also samples with the same values and proportions of ingredients, but with different compressive strength, probably due to differences in the building process. This is the case, for example, of samples 472, 473 and 474, shown in the table  \ref{tab:similar-samples-2}.

```{r similar-samples-2, echo=FALSE}
# Table - Same samples with different results
same_samples_2 <- dat %>%
  filter(id %in% c(472, 473, 474))
caption <- "Same samples with different results"
kable(
  same_samples_2[order(same_samples_2$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("scale_down"))
```

To facilitate the analysis of the samples, all samples that are the same in relation to the ingredients, have been assigned the same *id*. In addition, as the compressive strength at 28 days is the parameter to determine the grade of the concrete, only the elements containing that day among its samples were maintained. In the case of the same samples but with different results of compressive strength, the values were averaged. After all these changes (\ref{show-initial-data-cleaning}), the new total of samples was reduced to 970, containing 416 different settings for the proportions of ingredients.

```{r initial-data-cleaning, echo=FALSE}
# Data cleaning
dat <- dat %>%
  group_by(
    cement,
    blast_furnace_slag,
    fly_ash,
    water,
    superplasticizers,
    coarse_aggregate,
    fine_aggregate,
  ) %>%
  filter("28" %in% day) %>%
  mutate(id = id[which.min(id)]) %>%
  ungroup() %>%
  group_by(id, day) %>%
  mutate(mpa = mean(mpa)) %>%
  ungroup()
dat <- dat[!duplicated(select(dat, -c(id))),]
dat$id<-factor(dat$id)
n_samples <- nrow(dat)
n_distinct_samples <- n_distinct(dat$id)
```

The result can be seen in the table \ref{tab:similar-samples-same-id}. All samples with equal ingredient settings have the same id, and when they had different results for the same days, they were transformed into just one sample, with the arithmetic mean in the compressive strength.

```{r similar-samples-same-id, echo=FALSE, include=TRUE}
# Table - Previous samples after processing
same_samples <- dat %>%
  filter(id == 653 | id == 472 & day == 28)
caption <- "Previous samples after processing"
kable(
  same_samples[order(same_samples$id, same_samples$day),],
  col.names = dfUnits,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  row.names = FALSE,
  digits = 2
  ) %>%
  add_header_above(header = colNames, line = F, align = "c") %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

### Age selection

As previously described, the main age for analysis of compressive strength is 28 days, but other ages can also be used to build predictive models. However, it is necessary to verify how relevant the data of these other ages are. Starting with the distribution of the samples in relation to each age  (\ref{show-boxplot}) shown in the figure \ref{fig:boxplot}.

```{r boxplot, echo=F, fig.cap=cap, fig.height=3.5}
# Figure - Compressive strength (MPa) vs age (days)
cap <- "Boxplot - Compressive strength (MPa) vs age (days)"
ylabel <- "Compressive strength (MPa)"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x=factor(day), y=mpa)) +
  geom_boxplot() +
  geom_jitter(alpha=0.2)  +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel)
```

It was observed that the ages of 90, 91 and 100 days probably represent extremes to each other in the ingredient configurations, since they are relatively close ages but with very different values, especially for 90 and 91.

This hypothesis was verified using the principal component analysis method, applied to samples of these 3 ages (\ref{show-pca-90-91-100}). The figure \ref{fig:pca-90-91-100} shows how the samples relate to each other (which are similar or different) and revealed how each variable contributes to the analysis. The first two dimensions represent 37% and 24% $ respectively of the variance.

```{r pca-90-91-100, echo=F, include=T, fig.cap = cap, fig.height=3.5}
# Figure - Principal component analysis - 90, 91 e 100 days
dat_90_91_100 <- dat %>%
  ungroup() %>%
  filter(day %in% c(90, 91, 100)) %>%
  select(-c(id, mpa))
cap <- "Principal component analysis - 90, 91 e 100 days"
colnames(dat_90_91_100) <- c(
   "Cem.", "B.F.S.", "Fly.A.","Water","Sup.","C.Ag.","F.Ag.","day")
pca <- prcomp(select(dat_90_91_100, -c(day)), scale = TRUE)
habillage <- dat_90_91_100$day
fviz_pca_biplot(
  pca,
  geom.ind = "point",
  habillage=habillage,
  addEllipses = TRUE,
  ellipse.level=0.75) +
  ggtitle("") +
  theme_bw() +
  coord_cartesian(xlim = c(-3, 3.5), ylim = c(3, -5))
```

Another important point considered, of the concrete nature itself, is the fact that the growth rate of its compressive strength decreases with time, reaching a certain stability value. The figure \ref{fig:mpa-on-time} shows the compressive strength over the days for samples with more than 5 data, that is, data available for at least 6 different ages (\ref{show-mpa-on-time}).

```{r mpa-on-time, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Compressive strength through time
cap <- "Compressive strength through time"
ylabel <- "MPa"
xlabel <- "Age (days)"
dat_duplicated_only <- dat %>%
  group_by(id) %>%
  filter(n()>5) %>%
  select(id, mpa, day)
dat_duplicated_only %>%
  ggplot(aes(day, mpa, fill=id, alpha = 0.5)) +
  geom_line() +
  xlab(xlabel) +
  ylab(ylabel) +
  theme_bw() +
  theme(legend.position = "none")
```

For the reasons presented in the figures \ref{fig:boxplot}, \ref{fig:pca-90-91-100} and \ref{fig:mpa-on-time}, it was considered that the ages of 90, 91 and 100 days can be grouped to improve reading and decrease sample noise. They were converted to the same value, the age of 100 days was chosen (\ref{show-join-90-91-100}). As shown in the figure \ref{fig:mpa-on-time}, the resistance only increases, after 100 days the resistance to compression will be greater than or equal to the value of 90 or 91 days.

```{r join-90-91-100, echo=F}
# Joining 90, 91 and 100 days data
ind_90 <- dat$id[which(dat$day == "90")]
ind_91 <- dat$id[which(dat$day == "91")]
ind_100 <- dat$id[which(dat$day == "100")]
sum_duplicated <- sum(duplicated(c(ind_90, ind_91, ind_100))) # 0
dat <- dat %>%
  ungroup() %>%
  mutate(day = ifelse(day %in% c(91, 90), 100, day))
```

Another topic analyzed in the selection of ages was the observed frequency of each age value after this transformation from 90, 91 in 100 days, shown in the figure \ref{fig:freq-ages}. Some values of days have very low concentrations of samples, at the risk of damaging more than helping to create the models, so they were removed (\ref{show-remove-ages-lower-50}). The criterion adopted was to maintain only ages with a frequency greater than 50, only the values of 3, 7, 14, 28, 56 and 100 days.

```{r freq-ages, echo=F, include=T, fig.cap=cap, fig.height=3}
# Figure - Ages frequency
cap <- "Ages frequency"
ylabel <- "Frequency"
xlabel <- "Age (days)"
dat %>%
  ggplot(aes(x = factor(day))) +
  geom_bar() +
  theme_bw() +
  xlab(xlabel) +
  ylab(ylabel)
```

```{r remove-ages-lower-50, echo=F}
# Removing ages with frequency lower than 50
dat <- dat[dat$day %in% c(3, 7, 14, 28, 56, 100),]
```

### Data reorganization

The samples were grouped to maintain only one sample from each set of configuration of the proportions of the ingredients, adding new variables/columns for resistance at each age (\ref{show-reorganizing-dat}). The result in the first samples after this processing is shown in the table \ref{tab:new-features}.

```{r reorganizing-dat, echo=F}
# Data reorganization
dat <- dat %>%
  group_by_at(vars(-mpa)) %>%
  mutate(row_id = 1:n()) %>% ungroup() %>%
  spread(day, mpa, sep = "_") %>%
  select(-row_id)
```

```{r new-features, echo=F}
# Table - First 6 samples after reorganization
caption <- "First 6 samples after reorganization"
colNames2 = c("ID", "Cement", "B.F.S", "Fly ash", "Water",
             "Superp.", "Coarse Ag.", "Fine Ag.", "3 days", "7 days",
             "14 days", "28 days", "56 days", "100 days")
dfUnits2 <- c("", "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$kg/m^3$",
              "$kg/m^3$", "$kg/m^3$","$kg/m^3$","$MPa$","$MPa$",
              "$MPa$","$MPa$","$MPa$","$MPa$")
kable(
    head(dat[order(dat$id),]),
    col.names = dfUnits2,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c"
    ) %>%
    add_header_above(header = colNames2, line = F, align = "c") %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"))
```

The number of samples and distinct samples after all this manipulation remained the same, a total of 416 (\ref{show-total-samples-2}).

```{r total-samples-2, echo=F}
# Total samples
n_samples <- nrow(dat) # 416
n_distinct_samples <- n_distinct(dat$id) # 416
```

### Adding new variables

To finish the data preparation, new columns were added to the dataset (\ref{show-new-features-2}). Starting with the concrete class, for example if the compressive strength is between 25 and 30, it receives the class *C25*. The inclusion of the class was important because the compressive strength in *MPa* is a continuous variable, which will be used in the regression models, but the class as a discrete variable can provide another visualization of the data. The approximate mix of concrete was also added, which represents the proportions of aggregates (fine and coarse) for cement. Other proportions between the main ingredients were also added. The new variables are presented in the table \ref{tab:new-features-table}.

```{r new-features-2, echo=F}
# Adding New features
concrete_class <- function(mpa){
  if (mpa >= 10) {
    s <- as.character(mpa)
    first <- substr(s, start = 1, stop = 1)
    second <- ifelse(substr(s, start = 2, stop = 2) >= 5, 5, 0)
  }
  else {
    first <- ""
    second <- "5"
  }
  paste("C", first, second, sep = "")
}
mix <- function(c, f_ag, c_ag){
  paste(
    1,
    round(f_ag/c, 0),
    round(c_ag/c, 0),
    sep = ":")
}
dat <- dat %>%
  mutate(class = sapply(day_28, concrete_class)) %>%
  mutate(class = as.factor(class)) %>%
  mutate(mix_app = factor(
    mix(cement, fine_aggregate, coarse_aggregate))) %>%
  mutate(`water_/_cement` = water / cement) %>%
  mutate(`fine_aggregate_/_cement` = fine_aggregate/cement) %>%
  mutate(`coarse_aggregate_/_cement` = coarse_aggregate/cement) %>%
  mutate(`fine_aggregate_/_coarse_aggregate` = fine_aggregate/coarse_aggregate) %>%
  mutate(`water_/_coarse_aggregate` = water/coarse_aggregate) %>%
  mutate(`water_/_fine_aggregate` = water/fine_aggregate)
lvl <- levels(dat$class)
dat$class <- factor(
  dat$class, 
  levels=c( "C5", sort(lvl[lvl!="C5"], decreasing=F)))
```

```{r new-features-table, echo=F}
# Table - New features
caption <- "New features"
colNames7 = c("ID", "Class","Approximated Mix",
             "Water / Cement", "Fine Ag. / Cement", 
             "Coarse Ag. / Cement", "Fine Ag. / Coarse Ag.",
             "Water / Coarse Ag.", "Water / Fine Ag.")
kable(
    head(dat[order(dat$id),][,c(1,15:22)]),
    col.names = colNames7,
    escape = F,
    booktabs = T,
    caption = caption,
    linesep = "\\addlinespace",
    align = "c",
    digits = 4
    ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(2, width = "1.5cm") %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4:9, width = "1.7cm")
```

## Data visualization

In order to assess the need for further manipulation before building the models, in this step the 416 samples already processed were visualized and analyzed.

### Descriptive statistics

The table \ref{tab:stat-summ} presents the statistical data of the continuous variables (\ref {show-stat-summ}). The *Null* line represents the number of zeroed values for the ingredients, and the *NA* line represents the number of missing data. As the samples were filtered to maintain only sets of samples with known values of compressive strength at 28 days, the number of *NAs* is zero for that age. The figure \ref{fig:stat-summ-categorical} presents the statistical data of the discrete variables (\ref{show-stat-summ-categorical}).

```{r stat-summ, echo=FALSE}
# Table - Descriptive statistics - continuous variables
summ <- t(
  stat.desc(select(dat, -c(id, class, mix_app))))
caption <- "Descriptive statistics - continuous variables"
colnames(summ) <- c("Samples", "Null", "NA", "Min", "Max", "Range",
              "Sum", "Median", "Mean", "SE mean", 
              "CI mean","Variance", "Std.Dev.", "Coef.Var")
rownames(summ) = c("Cement", "B.F.S.", "Fly ash", "Water",
              "Superplast.", "Coarse agg.", "Fine agg.", "3 days", "7 days",
              "14 days", "28 days", "56 days", "100 days",
              "Water / Cement", "Fine agg. / Cement", 
              "Coarse agg. / Cement", "Fine agg. / Coarse agg.",
              "Water / Coarse agg.", "Water / Fine agg.")
kable(
  summ,
  escape = F,
  booktabs = T,
  caption = caption,
  linesep = "\\addlinespace",
  align = "c",
  digits = c(0,0,0,2,2,2,2,2,2,2,2,2,2,2)
  ) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down")) %>%
  column_spec(c(1,10:15), width = "1.5cm")
```

```{r stat-summ-categorical, echo=FALSE, fig.height=4, fig.cap = cap}
# Figure - Descriptive statistics - categorical variables
cap <- "Descriptive statistics - categorical variables"
name1 <- "Percentage"
name2 <- "Accumulated percentage"
ylabel <- "Frequency"
xlabel1 <- "Class"
xlabel2 <- "Approximate Mix"
format_percent = function(n){
  paste(n, "%", sep = "")
}
format_class <- function(cls){
  str_remove_all(cls, "C")
}
f_class <- freq(dat$class, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(class = row.names(.)) %>%
  mutate(class_n = as.numeric(format_class(class)))
f_cls_labels = function(n) {
  f_class$class[n]
}
f_cls_acc_labels = function(n){
  paste(f_class$`%cum`[n], "%", sep = "")
}
p1 <- f_class %>%
  ggplot(aes(x = as.integer(reorder(class_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$class) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_cls_labels, breaks = 1:16, limits = c(0.5,16.5),
                     sec.axis = sec_axis(~., breaks = 1:16,
                              name = name2,
                              labels = f_cls_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel1) +
  ylab(ylabel) +
  coord_flip()
format_mix <- function(mix){
  str_remove_all(mix, ":")
}
f_mix <- freq(dat$mix_app, cum = TRUE, sort = "dec", total = F) %>%
  select(n, "%", "%cum") %>%
  mutate(mix = row.names(.)) %>%
  mutate(mix_n = as.numeric(format_mix(mix)))
f_mix_labels = function(n) {
  f_mix$mix[n]
}
f_mix_acc_labels = function(n){
  paste(f_mix$`%cum`[n], "%", sep = "")
}
p2 <- f_mix %>%
  ggplot(aes(x = as.integer(reorder(mix_n, -n)), y = n)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(
      sec.axis = sec_axis(~./length(dat$mix_app) * 100,
                          name = name1,
                          labels = format_percent)) +
  scale_x_continuous(labels = f_mix_labels, breaks = 1:24, limits = c(0.5,24.5),
                     sec.axis = sec_axis(~., breaks = 1:24,
                              name = name2,
                              labels = f_mix_acc_labels)) +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  xlab(xlabel2) +
  ylab(ylabel) +
  coord_flip()
grid.arrange(p1, p2, ncol=2)
```

### Correlation between ingredients and compressive strength

The figure \ref{fig:correlation} shows the correlation of variables for each set of ages (\ref{show-correlation}). The figure \ref{fig:correlation-mpa} presents the same data, but instead of correlating them all, it only correlates with the compressive strength, showing the values in more detail (\ref{show-correlation-mpa}).

```{r correlation, echo=F, fig.height=4, fig.cap=cap}
# Figure - Correlation grouped by age
cor_dat <- dat %>% select(-c(id))
cap <- "Correlations at each age"
f_lvl <- c("3 days", "7 days", "14 days","28 days", "56 days", "100 days")
name <- "Correlation"
colnames_dat <- c("Cement", "B.F.S.", "Fly ash", "Water",
                "Superplast.", "Coarse agg.", "Fine agg.",
                "3", "7", "14", "28", "56", "100",
                "class", "mix_app", "W./C.", "F.A./C.", 
                "C.A./C.", "F.A./C.A.",
                "W./C.A.", "W./F.A.")
colnames(cor_dat) <- colnames_dat
cor_dat <- cor_dat %>%
  gather("day", "mpa", c("3", "7", "14", "28", "56", "100")) %>%
  drop_na()
cor_day <- function(d){
  res <- cor_dat %>%
    filter(day == d) %>%
    select(-c(day, class, mix_app)) %>%
    cor(.)
  res[upper.tri(res)] <- NA
  return(res)
}
cor_dats <- list(cor_day(3), cor_day(7), cor_day(14),
               cor_day(28), cor_day(56), cor_day(100))
melt_day <- function(df, d){
  df %>%
    melt() %>%
    mutate(day = d)
}
melt_dats <- list(melt_day(cor_dats[1], 3), melt_day(cor_dats[2], 7),
                  melt_day(cor_dats[3], 14), melt_day(cor_dats[4],28),
                  melt_day(cor_dats[5], 56), melt_day(cor_dats[6],100))
melt_dat_final <- melt_dats %>%
  reduce(rbind) %>%
  filter(value != 1)
melt_dat_final$day <- factor(melt_dat_final$day)
levels(melt_dat_final$day) <- f_lvl
melt_dat_final %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) + 
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=3) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
```

```{r correlation-mpa, echo=F, fig.height=3, fig.cap=cap}
# Figure - Correlation over time
cap <- "Correlation of variables with compressive strength over time"
name <- "Correlation"
melt_dat_final %>%
  filter(Var1 == "mpa" | Var2 == "mpa") %>%
  ggplot(aes(x=reorder(Var1, desc(Var1)), y=Var2, fill=value)) +
  geom_tile(color = "white") +
  facet_wrap(~day, ncol=6) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white",
   midpoint = 0, limit = c(-1,1),name=name, na.value="white") +
  xlab("") +
  ylab("") +
  geom_text(aes(label = round(value, 2)), size = 2.5) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))
```

The interpretation of the figure \ref{fig:correlation-mpa} suggests that the strength of the concrete is positively related mainly to the cement and superplasticizer ingredients and negatively to the water and fine aggregate. The smaller the amount of cement for aggregates and water, the more negatively they are correlated with compressive strength.

```{r mix-app-mpa, echo=F, fig.cap=cap,fig.height=3.5, warning=FALSE}
# Figure - Relationship between approximated mix, water, MPa and age
cap <- "Relationship between approximated mix, water, MPa and age"
d <- " days"
xlabel <- "Approximated mix" 
ylabel <- "Compressive strength (MPa)"
label <- "Water /\nCement"
mix_dat <- dat %>%
  select(c(day_3,day_7,day_14,day_28,day_56,day_100,
          mix_app, `water_/_cement`, 
          `fine_aggregate_/_cement`, 
          `coarse_aggregate_/_cement`))

labs <- paste(c("3","7","14","28","56","100"), d, sep="")
mix_dat <- mix_dat %>%
  gather("day", "mpa", -c(mix_app, `water_/_cement`,
                          `fine_aggregate_/_cement`,
                          `coarse_aggregate_/_cement`)) %>%
  drop_na()
lvls <- paste("day_",c("3","7","14","28","56","100"), sep="")
mix_dat$day <- factor(mix_dat$day, levels=lvls)
levels(mix_dat$day) <- labs
min_x <- min(mix_dat$`water_/_cement`)
max_x <- max(mix_dat$`water_/_cement`)
s_x <- max_x - min_x
mix_dat %>%
  ggplot(aes(x=mix_app, y=mpa, colour = `water_/_cement`)) +
  geom_point()  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_shape_manual(values=c(16, 2, 8)) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = s_x / 2 + min_x ,limits = c(min_x, max_x),
                         breaks = c(round(min_x, 2),
                                    round(s_x*0.25 + min_x,2), 
                                    round(s_x*0.5 + min_x,2),
                                    round(s_x * 0.75 + min_x,2),
                                    round(max_x, 2))) +
  labs(colour = label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```

```{r mix-mpa, echo=F, fig.cap=cap, fig.height=3.5, warning=FALSE, message=F}
# Figure - Relationship between concrete main features
cap <- "Relationship between concrete main features"
d <- " days"
xlabel <- "Water / Cement"
ylabel <- "Compressive strength (MPa)"
colour <- "Fine Agg. /\nCement"
size <- "Coarse Agg. /\nCement"
min_x_2 <- min(mix_dat$`fine_aggregate_/_cement`)
max_x_2 <- max(mix_dat$`fine_aggregate_/_cement`)
s_x_2 <- max_x_2 - min_x_2
mix_dat %>%
  ggplot(aes(x=`water_/_cement`, y=mpa,
             colour = `fine_aggregate_/_cement`,
             size = `coarse_aggregate_/_cement`)) +
  geom_point(alpha = 0.5)  +
  facet_wrap(~ day, ncol=2) +
  theme_bw() +
  ylab(ylabel) +
  xlab(xlabel) +
  scale_colour_gradient2(low = "red", mid = "yellow", high = "blue",
                         midpoint = (s_x_2 / 2) + min_x_2, 
                         limits = c(min_x_2, max_x_2),
                         breaks = c(round(min_x_2, 2),
                                    round(s_x_2*0.25 + min_x_2,2),
                                    round(s_x_2*0.5 + min_x_2,2),
                                    round(s_x_2 * 0.75 + min_x_2,2),
                                    round(max_x_2 - 0.01, 2))
                         ) +
  labs(colour = colour, size = size) +
  ylim(c(-5,85)) +
  scale_radius() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"))
```

The figures \ref{fig:mix-app-mpa} and \ref{fig:mix-mpa} show the relationship between the main ingredients (known as mix) in relation to the compressive strength (\ref{show-mix-app-mpa} and \ref{show-mix-mpa}). The interpretation of these figures shows that the greater the amount of cement in relation to the other ingredients, the greater the resistance to compression.

### Variables distribution

>>>>

<EN>

@@

</EN>

<PT>

# Introdução

A resistência à compressão é a principal característica do concreto, medida por testes de padrões internacionais que consistem na quebra de corpos de prova. A medição aos 28 dias é obrigatória e representa a classe do concreto. Saber com antecêdencia qual o resultado será obtido para uma determinada idade, a partir das proporções de seus ingredientes, é de grande interesse para os fabricantes de concreto, construtoras e engenheiros civis.

@@

Essa resistência à compressão é uma função não linear de seus ingredientes e idade, tornando difícil o estabelecimento de uma fórmula analítica, apesar de algumas fórmulas já haveram sido propostas. @Hasan2011 propós um modelo matemático para prever a partir dos resultados de testes de 7 e 14 dias, e @Kabir2012 a partir de 7 dias. Porém técnicas de machine learning podem ser utilizadas para modelar essa característica a partir de dados reais de amostras, utilizando apenas os ingredientes.

@@

Muitos estudos anteriores utilizam o mesmo conjunto de dados utilizado por @Yeh1998 para prever a resistência à compressão do concreto. @Alshamiri2020 obteve bons resultados com a técnica de regularized extreme learning machine (RELM), e @Hameed2020 obteve resultados ainda melhores com a técnica de Artificial Neural Networks e cross-validation. Esse conjunto de amostras é tão conhecido que há ainda páginas na internet de estudos não publicados que o utilizam e possuem bons resultados, como @Abban2016, @Raj2018, @Modukuru2020 e @Pierobon2018. Ao final do trabalho os resultados encontrados são comparados aos trabalhos citados aqui.

@@

Diferente dos estudos anteriores com esse conjunto de amostras, este trabalho faz a preparação dos dados de forma diferente. A idade do concreto é a variável mais singular que contribui para sua resistência à compressão, por esse motivo, a idade é tratada separadamente nos modelos de machine learning, criando modelos para cada faixa de idade.

</PT>
